Hybrid platforms and bargaining power

### 2024-05-09 src/paper/application.tex
Rewrite second extension

Let us start by deriving total surplus as a function of $N_P, N_F$ and $N_C$.
Under the logit-like demand structure described in \cref{sec:example}, the total surplus is given by
\begin{proposition}
    \label{prop:profits_total_surplus}
    If the platform can extract consumer surplus, then the total surplus is given by
    \begin{align*}
        \Pi(N_P, N_F, N_C) = \mu N_C \left[ \frac{N_P V_P + N_F V_F}{N_P V_P + N_F V_F + 1} + \log(N_P V_P + N_F V_F + 1) \right].
    \end{align*}
\end{proposition}
\begin{proof}[Proof of \cref{prop:profits_total_surplus}]
    See \textcite{small1981applied} for a derivation for the discrete case.
    The continuous case is analogous.
\end{proof}
Note, that this expression is of the form $\Pi(N_P, N_P, N_C) = N_C f(N_P, N_F)$.
Furthermore, $f(N_P, N_F) = g(V_P N_P + V_F N_F)$ for an increasing, strictly concave $g$.

Next, let us consider the bargaining outcomes in this three-sided setting.
\Cref{sec:cooperative_game_two_sided} formally describes the corresponding cooperative game, and \cref{prop:profit_sharing_two_sided} establishes the resulting profit shares.
To summarize, the shares of the various players are given by
Then, the resulting profit shares are given by
\begin{align*}
    \pi_P(N_P, N_F, N_C) &= N_C \int_0^1 s f(N_P, s N_F) \ds, \\
    \pi_F(N_P, N_F, N_C) &= N_C \int_0^1 s^2 N_F \partial_2 f(N_P, s N_F) \ds \\
    CS(N_P, N_F, N_C) &= N_C \int_0^1 s f(N_P, s N_F) \ds.
\end{align*}

There are a number of things to note here.
First, and most importantly, the consumers' share from total surplus is equal to the platform's.\footnote{
    This is a consequence of the fact that the total surplus function is linear in the number of consumers.
}
As a consequence, what's good for the platform is also optimal for the consumers.
Therefore, letting the platform choose the number of its products maximizes consumer welfare, as well (but not necessarily total surplus).

Second, the expression for the fringe profits has the same structure as in \cref{ass:profit_sharing}.
Therefore, if we show that \cref{ass:single_crossing} holds, it in turn implies that all of the results in \cref{sec:example} apply here, as well.
The following lemma establishes this result.
\begin{lemma}
    \label{lem:profit_assumptions_two_sided}
    $f(n_P, n_F)$ satisfies \cref{ass:single_crossing}.
    Furthermore,
    \begin{align*}
        \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} < 0.
    \end{align*}
\end{lemma}

With this proven, we can again rely on \cref{prop:aggregate_size_additive} to see that, in equilibrium, an increase in $N_P$ decreases $N_F$.
Furthermore, this decrease is so large that total product variety, and thus total surplus (which in turn is equal to total profits) also decreases (\cref{fig:fringe_entry_eq_full_surplus_two_sided}).\footnote{
    As long as the market is in the hybrid regime.
    If $N_F^* = 0$, then an increase in $N_P$ mechanically increases total aggregate, and thus total surplus.
}

Nevertheless, there is an important distinction compared to the previous results.
As before, platform profits are again increasing in $N_P$.
However, in this case, it also implies that consumers' share is also increasing in this case (\cref{fig:profits_full_surplus_two_sided}).
Therefore, consumers do prefer hybrid operation to the the platform being a pure marketplace.

>>>>

Let us start by deriving total surplus as a function of $N_P, N_F$ and $N_C$.
Under the logit-like demand structure described in \cref{sec:demand}, the total surplus is given by
\begin{proposition}
    \label{prop:profits_total_surplus}
    Assuming profit-maximizing prices from the platform and the fringe, total surplus as a function of $N_P, N_F$ and $N_C$ is given by
    \begin{align*}
        \Pi(N_P, N_F, N_C) = \mu N_C \left[ \frac{N_P V_P + N_F V_F}{N_P V_P + N_F V_F + 1} + \log(N_P V_P + N_F V_F + 1) \right].
    \end{align*}
\end{proposition}
\begin{proof}[Proof of \cref{prop:profits_total_surplus}]
    See \textcite{small1981applied} for a derivation for the discrete case.
    The continuous case is analogous.
\end{proof}
Note, that this expression is of the form $\Pi(N_P, N_P, N_C) = N_C f(N_P, N_F)$.
Furthermore, $f(N_P, N_F) = g(V_P N_P + V_F N_F)$ for an increasing, strictly concave $g$.

Next, let us consider the bargaining outcomes in this three-sided setting.
\Cref{sec:cooperative_game_two_sided} formally describes the corresponding cooperative game, and \cref{prop:profit_sharing_two_sided} establishes the resulting profit shares.
To summarize, the shares of the various players are as follows.
\begin{proposition}
    \label{prop:profit_sharing_two_sided}
    \begin{align*}
        \pi_P(N_P, N_F, N_C) &= \int_0^1 s \Pi(N_P, s N_F) \ds, \\
        \pi_F(N_P, N_F, N_C) &= \int_0^1 s^2 N_F \partial_2 \Pi(N_P, s N_F) \ds \\
        CS(N_P, N_F, N_C) &= \int_0^1 s \Pi(N_P, s N_F) \ds.
    \end{align*}
\end{proposition}

There are a number of things to note here.
First, and most importantly, the consumers' share from total surplus is equal to the platform's.\footnote{
    This is a consequence of the fact that the total surplus function is linear in the number of consumers.
}
As a consequence, what's good for the platform is also optimal for the consumers.
Therefore, if the platform can choose its mode of operation to maximize its profits, it maximizes consumer welfare, as well (but not necessarily total surplus).

Second, it can be shown that due to the fringe firms' total profit function having a similar shape as in the main model, the same results hold in terms of the platform's product variety displacing fringe products.
\begin{proposition}
    In the hybrid regime under three-sided bargaining, the following holds:
    \begin{align*}
        N_F > 0 \implies \frac{\partial N_F}{\partial N_P} < -\frac{V_P}{V_F}.
    \end{align*}
\end{proposition}
As a consequence, the aggregate is decreasing in $N_P$ in the hybrid regime (\cref{fig:fringe_entry_eq_full_surplus_two_sided}).
\begin{corollary}
    In the hybrid regime under three-sided bargaining, total aggregate is decreasing in the platform's product variety:
    \begin{align*}
        N_F > 0 \implies \frac{\partial A}{\partial N_P} < 0.
    \end{align*}
\end{corollary}

Nevertheless, there is an important distinction compared to the previous results.
While this does lead to a decrease in total surplus, in contrast to the two-sided bargaining case, this does not imply a decrese in consumer surplus.
As mentioned before, the consumers' share of total surplus is equal to the platform's profits, and thus it is increasing in $N_P$ if and only if platform profits (\cref{fig:profits_full_surplus_two_sided}) are.
Therefore, consumers might prefer hybrid operation to the platform being a pure marketplace.
